---
title: "Ardag et al. 2019 Representation Replication Material"
author: "M. Murat Ardag"
date: "9/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Visit [team populism](https://populism.byu.edu/){target="_blank"} for various other publications, data, resources and more.

Data & R markdown file also available at [https://osf.io/g8d53](https://osf.io/g8d53){target="_blank"} 

# load the necessary packages

```{r}
packages <- c("lavaan" , "semPlot", "semTools", "psych", "tidyverse", "reshape2", "knitr")

for(i in packages){
  if(!require(i, character.only = T)) install.packages(i, dependencies=T)
  library(i, character.only = T)
}
options(knitr.kable.NA = '')
```

# load the datasets

```{r}
load("dataUS.RData")
load("dataDE1.RData")
load("dataDE2.RData")
```

# Study 1

popexp1 = EXPTRT

convert the experimental treatment into factor

```{r}
US$EXPTRT <- factor(US$EXPTRT, levels = c(0,1), labels = c("Control","Treatment"))
```

## (+) political engagement strategy: active support in elections

```{r}
US_model_PE <- '
pop =~ ant1 + ant2 + ant3 + ant5 + pop2 + pop3 + pop4 # populist attitudes
ase =~ anes617 + anes616 + anes615 # active support in elections
ase ~ pop # regression
'
US_PE_r_fit <- sem (US_model_PE, US, estimator="mlr", mimic="mplus", missing="fiml",
                    group = "EXPTRT", sample.mean = T,
                    group.equal=c("loadings", "intercepts",
                                  "residuals", "residual.covariances", "lv.variances", "lv.covariances",
                                  "regressions"),
                    group.partial=c("means")) # restricted model

US_PE_f_fit <- sem (US_model_PE, US, estimator="mlr", mimic="mplus", missing="fiml",
                    group = "EXPTRT", sample.mean = T,
                    group.equal=c("loadings", "intercepts", 
                                  "residuals", "residual.covariances", "lv.variances", "lv.covariances"),
                    group.partial=c("regressions", "means")) # free model
```

## model comparison & summary

```{r}
anova(US_PE_r_fit, US_PE_f_fit)
cf <- compareFit(US_PE_r_fit, US_PE_f_fit, nested=T)
summary(cf)
```

## SEM estimates summary

```{r}
summary(US_PE_r_fit, fit.measures=T, standardized=T, rsquare=T)
summary(US_PE_f_fit, fit.measures=T, standardized=T, rsquare=T)
```

## plot the path models

```{r, fig.width=15, fig.height=8}
par(mfrow=c(2,2))
semPaths(US_PE_r_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
semPaths(US_PE_f_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
dev.off()
```

## (-) political engagement strategy: legitimate radical political action

```{r}
US_model_NE <- '
pop =~ ant1 + ant2 + ant3 + ant5 + pop2 + pop3 + pop4 # populist attitudes
rpa =~ radact1 + radact2 + radact3 + radact5 + radact6 # legitimate radical political action
rpa ~ pop # regression
'
US_NE_r_fit <- sem (US_model_NE, US, estimator="mlr", mimic="mplus", missing="fiml",
                    group = "EXPTRT", sample.mean = T,
                    group.equal=c("loadings", "intercepts",
                                  "residuals", "residual.covariances", "lv.variances", "lv.covariances",
                                  "regressions"),
                    group.partial=c("means"))

US_NE_f_fit <- sem (US_model_NE, US, estimator="mlr", mimic="mplus", missing="fiml",
                    group = "EXPTRT", sample.mean = T,
                    group.equal=c("loadings", "intercepts", 
                                  "residuals", "residual.covariances", "lv.variances", "lv.covariances"),
                    group.partial=c("regressions", "means"))
```

## model comparison & summary

```{r}
anova(US_NE_r_fit, US_NE_f_fit)
cf <- compareFit(US_NE_r_fit, US_NE_f_fit, nested=T)
summary(cf)
```

## SEM estimates summary

```{r}
summary(US_NE_r_fit, fit.measures=T, standardized=T, rsquare=T)
summary(US_NE_f_fit, fit.measures=T, standardized=T, rsquare=T)
```

## plot the path models

```{r, fig.width=15, fig.height=8}
par(mfrow=c(2,2))
semPaths(US_NE_r_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
semPaths(US_NE_f_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
dev.off()
```

# Study 2 DE sample 1

convert the experimental treatment into factor

```{r}
DE1$EXPTRT <- factor(DE1$EXPTRT, levels = c(0,3), labels = c("Control","Treatment"))
```

## measurement model
## get the single populism score from multiple dimensions

as explained in [Castanho Silva, Bruno, Ioannis Andreadis, Eva Anduiza, Nebojsa Blanusa, Yazmin Morlet Corti, Gisela Delfino, Guillem Rico, Saskia P. Ruth, Bram Spruyt, Marco Steenbergen, and Levente Littvay. 2018. "Public Opinion Surveys: a New Scale". In: The Ideational Approach to Populism: Theory, Method & Analysis, edited by Kirk A. Hawkins, Ryan Carlin, Levente Littvay, and Cristóbal Rovira Kaltwasser. London: Routledge. 150-177](https://www.routledge.com/The-Ideational-Approach-to-Populism-Concept-Theory-and-Analysis-1st/Hawkins-Carlin-Littvay-Kaltwasser/p/book/9781138716537){target="_blank"}

see also [Castanho Silva, Bruno, Sebastian Jungkunz, Marc Helbling, and Levente Littvay. 2019. "An Empirical Comparison of Seven Populist Attitudes Scales". Political Research Quarterly.](https://journals.sagepub.com/doi/10.1177/1065912919833176){target="_blank"}

```{r}
DE_S1_measure.model <-'
pc =~ POPpc3 + POPpc1
ae =~ POPae3 + POPae1
mw =~ POPmwv3 + POPmwv1
'
measure.cfa.fit <-cfa(DE_S1_measure.model, DE1, estimator='mlr', mimic='mplus',missing='fiml', sample.mean=T)
range01 <- function(x){(x-min(x,na.rm=T))/(max(x,na.rm=T)-min(x,na.rm=T))}
dimensions<-predict(measure.cfa.fit)[,1:3]
dimensions<-apply(dimensions,2,range01)
pop_DE1<-dimensions[,1]*dimensions[,2]*dimensions[,3]
DE1$pop<-pop_DE1
```

## (+) political engagement strategy: being an active citizen

```{r}
DE1_model_PE <-'
gc =~ alwVote + KWAoGov + ActSocPol
gc ~ pop # regression
'

DE1_PE_r_fit <- sem (DE1_model_PE,DE1, estimator="mlr", mimic="mplus", missing="fiml",
                     group = "EXPTRT", sample.mean = T,
                     group.equal=c("loadings", "intercepts",
                                   "residuals", "residual.covariances", "lv.variances", "lv.covariances",
                                   "regressions"),
                     group.partial=c("means"))

DE1_PE_f_fit <- sem (DE1_model_PE,DE1, estimator="mlr", mimic="mplus", missing="fiml",
                     group = "EXPTRT", sample.mean = T,
                     group.equal=c("loadings", "intercepts", 
                                   "residuals", "residual.covariances", "lv.variances", "lv.covariances"),
                     group.partial=c("regressions", "means"))
```


## model comparison & summary

```{r}
anova(DE1_PE_r_fit, DE1_PE_f_fit)
cf <- compareFit(DE1_PE_r_fit, DE1_PE_f_fit, nested=T)
summary(cf)
```

## SEM estimates summary

```{r}
summary(DE1_PE_r_fit, fit.measures=T, standardized=T, rsquare=T)
summary(DE1_PE_f_fit, fit.measures=T, standardized=T, rsquare=T)
```

## plot the path models

```{r, fig.width=15, fig.height=8}
par(mfrow=c(2,2))
semPaths(DE1_PE_r_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
semPaths(DE1_PE_f_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
dev.off()
```

## (-) political engagement strategy: abstaining from voting

```{r}
DE1_model_NE <-'
vot =~ a*vp1 + a*vp3 # not voting
vot ~ pop # regression
'
DE1_NE_r_fit <- sem (DE1_model_NE, DE1, estimator="mlr", mimic="mplus", missing="fiml",
                     group = "EXPTRT", sample.mean = T,
                     group.equal=c("loadings", "intercepts",
                                   "residuals", "residual.covariances", "lv.variances", "lv.covariances",
                                   "regressions"),
                     group.partial=c("means"))

DE1_NE_f_fit <- sem (DE1_model_NE, DE1, estimator="mlr", mimic="mplus", missing="fiml",
                     group = "EXPTRT", sample.mean = T,
                     group.equal=c("loadings", "intercepts", 
                                   "residuals", "residual.covariances", "lv.variances", "lv.covariances"),
                     group.partial=c("regressions", "means"))
```

## model comparison & summary

```{r}
anova(DE1_NE_r_fit, DE1_NE_f_fit)
cf <- compareFit(DE1_NE_r_fit, DE1_NE_f_fit, nested=T)
summary(cf)
```

## SEM estimates summary

```{r}
summary(DE1_NE_r_fit, fit.measures=T, standardized=T, rsquare=T)
summary(DE1_NE_f_fit, fit.measures=T, standardized=T, rsquare=T)
```

## plot the path models

```{r, fig.width=15, fig.height=8}
par(mfrow=c(2,2))
semPaths(DE1_NE_r_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
semPaths(DE1_NE_f_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
dev.off()
```

# Study 3 DE sample 2

convert into factor

```{r}
DE2$EXPTRT <- factor(DE2$EXPTRT, levels = c(0,1,2), labels = c("Control","PEGIDA","LEIPZIG"))
```

## measurement model
## get the single populism score from multiple dimensions

```{r}
DE2_measure.model <- '
pc =~ POPpc1 + POPpc2 + POPpc3 # people-centrism
ae =~ POPae1 + POPae2 + POPae3 # anti-elitism
mw =~ POPmw1 + POPmw2 + POPmw3 # manichean view of politics
'
DE2_measure.cfa <- cfa(DE2_measure.model, DE2, estimator='mlr', mimic='mplus',missing='fiml', sample.mean=T)
range01 <- function(x){(x-min(x,na.rm=T))/(max(x,na.rm=T)-min(x,na.rm=T))}
dimensions<-predict(DE2_measure.cfa)[,1:3]
dimensions<-apply(dimensions,2,range01)
pop_DE2 <- dimensions[,1]*dimensions[,2]*dimensions[,3]
DE2$pop<-pop_DE2
```

## political engagement strategy:
## (+) engagement for the support of the Leipzig demonstration
## (-) engagement for the support of the PEGIDA demonstration

```{r}
DE2_model <- '
dsim =~ GrpIDdsim1 + GrpIDdsim2 + GrpIDdsim3 # group dissimilarity
datt =~ GrpIDdatt1 + GrpIDdatt2 + GrpIDdatt3 # group detachment
dsim ~~ datt # correlation between endogeneous variables
dsim ~ pop # regression
datt ~ pop # regression
'
DE2_r_fit <- sem (DE2_model,DE2, estimator="mlr", mimic="mplus", missing="fiml",
                  group = "EXPTRT", sample.mean = T,
                  group.equal=c("loadings", "intercepts",
                                "residuals", "residual.covariances", "lv.variances", "lv.covariances",
                                "regressions"),
                  group.partial=c("means"))

DE2_f_fit <- sem (DE2_model,DE2, estimator="mlr", mimic="mplus", missing="fiml",
                  group = "EXPTRT", sample.mean = T,
                  group.equal=c("residuals", "residual.covariances", "lv.variances", "lv.covariances"),
                  group.partial=c("loadings", "intercepts", "regressions", "means"))
```

## model comparison & summary

```{r}
anova(DE2_r_fit, DE2_f_fit)
cf <- compareFit(DE2_r_fit, DE2_f_fit, nested=T)
summary(cf)
```

## SEM estimates summary

```{r}
summary(DE2_r_fit, fit.measures=T, standardized=T, rsquare=T)
summary(DE2_f_fit, fit.measures=T, standardized=T, rsquare=T)
```

## plot the path models

```{r, fig.width=15, fig.height=8}
par(mfrow=c(2,2))
semPaths(DE2_r_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
semPaths(DE2_f_fit, "mod", "std", intercept=F, rotation = 2, edge.label.cex=.9, sizeMan = 5, sizeLat = 6, optimizeLatRes=T, ask=F)
dev.off()
```

# measurement models & construct reliabilities

## US

```{r, fig.width=15, fig.height=8}
model <- '
pop =~ ant1 + ant2 + ant3 + ant5 + pop2 + pop3 + pop4 # populist attitudes
ase =~ anes617 + anes616 + anes615 # active support in elections
rpa =~ radact1 + radact2 + radact3 + radact5 + radact6 # legitimate radical political action
'
fit <- cfa (model, data=US, estimator="mlr", mimic="mplus", missing="fiml", std.ov=T, std.lv=T)
summary(fit, fit.measures=T, standardized=T, rsquare=T)
semPaths(fit, "mod", "std", intercepts = F, edge.label.cex=1, rotation = 4, layout = "tree3")
round(reliability(fit),2)
```

## DE1

```{r, fig.width=15, fig.height=8}
model <- '
pc =~ POPpc3 + POPpc1 # people-centrism
ae =~ POPae3 + POPae1 # anti-elitism
mw =~ POPmwv3 + POPmwv1 # manichean view of politics
gc =~ alwVote + KWAoGov + ActSocPol # active citizen (+) engagement
vot =~ vp3 + vp1 # not voting (-) engagement
'
fit <- cfa (model, data=DE1, estimator="mlr", mimic="mplus", missing="fiml", std.ov=T, std.lv=T)
summary(fit, fit.measures=T, standardized=T, rsquare=T)
semPaths(fit, "mod", "std", intercepts = F, edge.label.cex=1, rotation = 4, layout = "tree3")
round(reliability(fit),2)
```

## DE2

```{r, fig.width=15, fig.height=8}
model <- '
pc =~ POPpc1 + POPpc2 + POPpc3 # people-centrism
ae =~ POPae1 + POPae2 + POPae3 # anti-elitism
mw =~ POPmw1 + POPmw2 + POPmw3 # manichean view of politics
dsim =~ GrpIDdsim1 + GrpIDdsim2 + GrpIDdsim3 # group dissimilarity
datt =~ GrpIDdatt1 + GrpIDdatt2 + GrpIDdatt3 # group detachment
'
fit <- cfa (model, data=DE2, estimator="mlr", mimic="mplus", missing="fiml", std.ov=T, std.lv=T)
summary(fit, fit.measures=T, standardized=T, rsquare=T)
semPaths(fit, "mod", "std", intercepts = F, edge.label.cex=1, rotation = 4, layout = "tree3")
round(reliability(fit),2)
```

# get the descriptive statistics

## US

```{r}
US %>% 
  select(ant1,ant2,ant3,ant5,pop2,pop3, pop4,anes617,anes616,anes615,radact1,radact2,radact3,radact5,radact6) %>% 
  gather("Variable", "value") %>% 
  group_by(Variable) %>% 
  summarise(Mean=mean(value, na.rm=T), 
            SD=sd(value, na.rm=T), 
            min=min(value, na.rm=T), 
            max=max(value, na.rm=T), 
            '% Missing'=100*length(which(is.na(value)))/n()) %>% 
  kable(digits=2, format="pandoc", caption="Descriptive Statistics US")
US_fdp <- US %>% select(ant1,ant2,ant3,ant5,pop2,pop3, pop4,anes617,anes616,anes615,radact1,radact2,radact3,radact5,radact6)
s1dp <- ggplot(melt(US_fdp),aes(x=value)) + geom_density() + facet_wrap(~variable) + ggtitle("US: Density plots of the items used in Study 1")
```

## DE1

```{r}
DE1 %>% 
  select(POPpc3,POPpc1,POPae3,POPae1,POPmwv3,POPmwv1,alwVote,KWAoGov,ActSocPol,vp3,vp1) %>% 
  gather("Variable", "value") %>% 
  group_by(Variable) %>% 
  summarise(Mean=mean(value, na.rm=T), 
            SD=sd(value, na.rm=T), 
            min=min(value, na.rm=T), 
            max=max(value, na.rm=T), 
            '% Missing'=100*length(which(is.na(value)))/n()) %>% 
  kable(digits=2, format="pandoc", caption="Descriptive Statistics DE 1")
DE1_fdp <- DE1 %>% select(POPpc3,POPpc1,POPae3,POPae1,POPmwv3,POPmwv1,alwVote,KWAoGov,ActSocPol,vp3,vp1)
s2dp <- ggplot(melt(DE1_fdp),aes(x=value)) + geom_density() + facet_wrap(~variable) + ggtitle("DE: Density plots of the items used in Study 2")
```

## DE2

```{r}
DE2 %>% 
  select(POPpc1,POPpc2,POPpc3,POPae1,POPae2,POPae3,POPmw1,POPmw2,POPmw3,GrpIDdsim1,GrpIDdsim2,GrpIDdsim3,GrpIDdatt1,GrpIDdatt2,GrpIDdatt3) %>% 
  gather("Variable", "value") %>% 
  group_by(Variable) %>% 
  summarise(Mean=mean(value, na.rm=T), 
            SD=sd(value, na.rm=T), 
            min=min(value, na.rm=T), 
            max=max(value, na.rm=T), 
            '% Missing'=100*length(which(is.na(value)))/n()) %>% 
  kable(digits=2, format="pandoc", caption="Descriptive Statistics DE2")
DE2_fdp <- DE2 %>% select(POPpc1,POPpc2,POPpc3,POPae1,POPae2,POPae3,POPmw1,POPmw2,POPmw3,GrpIDdsim1,GrpIDdsim2,GrpIDdsim3,GrpIDdatt1,GrpIDdatt2,GrpIDdatt3)
s3dp <- ggplot(melt(DE2_fdp),aes(x=value)) + geom_density() + facet_wrap(~variable) + ggtitle("DE: Density plots of the items used in Study 3")
```


```{r, fig.width=15, fig.height=8}
gridExtra::grid.arrange(s1dp,s2dp,s3dp, nrow=2, ncol=2)
```


# Study 2
## Figure 1: self-report political orientation

```{r}
table(DE1$POlr)
DE1$POlr <- factor(DE1$POlr,
                   levels = c(1,2,3,4,5,6,7,8),
                   labels = c("Left","2","3", "4","5","6","7","Right"))

DE1 %>% select(POlr) %>% drop_na() %>% ggplot(., aes(x=POlr)) + 
  geom_bar(aes(y = (..count..)),fill="grey", alpha=0.6) + 
  geom_text(aes(y = (..count..) ,label =   ifelse((..count..)==0,"",scales::percent((..count..)/sum(..count..)))), stat="count",colour="black") +
  theme_bw() +
  xlab("Self-Report Political Orientation in %") +
  ylab("Frequency") + 
  labs(title="Political orientation of the German university students in study 2")
```


# Study 3
## Figure 2: self-report political orientation

```{r}
table(DE2$PolOrLR)
DE2$PolOrLR <- factor(DE2$PolOrLR,
                      levels = c(1,2,3,4,5,6,7),
                      labels = c("Left","2","3", "4","5","6","Right"))
DE2 %>% select(PolOrLR) %>% drop_na() %>% ggplot(., aes(x=PolOrLR)) + 
  geom_bar(aes(y = (..count..)),fill="grey", alpha=0.6) + 
  geom_text(aes(y = (..count..) ,label =   ifelse((..count..)==0,"",scales::percent((..count..)/sum(..count..)))), stat="count",colour="black") +
  theme_bw() +
  xlab("Self-Report Political Orientation in %") +
  ylab("Frequency") + 
  labs(title="Political orientation of the German university students in study 3")
```

